---
title: "WUP_04"
author: "April Parry"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
require(stats)
thedata=read_csv("GSScsv.csv")
```


```{r eval=FALSE}
thedata<-thedata %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We wanted to find out by looking at the data what sort of data we had. We copied write up 3 to read the data in and changed the character variables into factor variables and saw that we ad a numerical data which was WORDSUM and we now have a factor variable called DEGREE. So we deduced that we have a numerical~ categorical problem. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}

# Creating the fancy box plot
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = NA) +
  theme_minimal() +
  labs(x = "DEGREE", y = "WORDSUM") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Fancy Box Plot")




```

LT High School would not overlap with junior College or Graduate, Thus showing that it is very likely to be different schooling. In addition, Graduate does not overlap with High School. Therefore, it also likely to be different. Ultimately, LT High School and Graduate are not only different from one another they are both different that another type of schooling, such as High School and Junior College. 


### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```

The average score on the WORDSUM test for graduates was a little over 10. There's a difference between the average score in Graduate school and the average score in junior College. Thus further it provides that the schooling are different than each other. 

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=thedata)
summary.aov(model1)
```

There is a significant difference between the schooling of all 5 degrees.

```{r}

# Performing ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Performing Tukey's HSD test
tukey_result <- TukeyHSD(model)

# Viewing the test results
tukey_result

```

Based on the chart above, LT High School and Graduate are the two degrees that are different from one another. In addition, High School is also different than Graduate. 


```{r}
# Assuming you have the "thedata" data frame with "DEGREE" and "WORDSUM" columns
# Load the emeans package
library(emmeans)

# Perform one-way ANOVA
result <- aov(WORDSUM~DEGREE, data = thedata)

# perform Scheffe's post hoc test
comparison <- emmeans(result, pairwise ~ DEGREE, adjust = "scheffe")

# Print the comparison table
print(comparison)
```

```{r}
# Performing ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Installing and loading the 'agricolae' package
install.packages("agricolae")
library(agricolae)

# Performing Scheffé's test for pairwise comparisons
scheffe_result <- LSD.test(model, "DEGREE", alpha = 0.05)

# Viewing the test results
scheffe_result
```



<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The two different categories would be the LT High School graduate and the High School graduates since both of them don't run into the zero line. These two categories are the only two that would fall outside of that which we can see in the graph.  
